---
title: "fifty percent reduction"
output: html_document
---

```{r setup, include=FALSE}
library(readxl)
library(data.table)
library(dplyr)
library(tidyr)
library(naniar)
library(caret)
"%&%" = function(a,b) paste0(a,b)
set.seed(1234)
```


```{r}

data.dir<-"C:\\Users\\rshoo\\OneDrive\\Desktop\\rush\\ML project\\datasets\\3 Week IOP\\"
list.files(data.dir)
ThreeWeekData<-read_xlsx(data.dir %&% "IOP_DATA_SET_6-12-2020 with Manual Corrections.xlsx",na=c("N/A","NULL","Not Applicable","NA","\\s","*Unknown","*Unspecified","Unknown","Refusal","No Response","-", "Not Available"))
out.dir<-"C:\\Users\\rshoo\\OneDrive\\Desktop\\rush\\ML project\\replicate all response models\\"
```

get relevant variable names from reliable change data


```{r}
dir<-"C:\\Users\\rshoo\\OneDrive\\Desktop\\rush\\ML project\\reliable change\\"
RCIdata<-fread(dir %&% "reliablechangedataset.csv")
colnames(RCIdata)
# rm(RCIdata)
```

demographic variables
"sex", x
"age", x
"post911", x
"ethnic", x
"race", x
"deployed", x
"nsivalid", x
"nsi", x
"mst" x
one hot encoded demographic variables
referral, - not in data set
marriage, x
branch, x
education x

baseline psychiatric and medical variables
caps scores x
ptci x
phq x
pcl x
"auditc"
"anxietytx - not in dataset
"depressiontx" - not in dataset
"substancetx"- not in dataset
"ptsdtx" - not in dataset
"tbitx" - not in dataset

day 14 PCL scores x


```{r}
# 
#                      "InjuryDescription_4",
#                      "InjuryDescription_8",
#                      "InjuryDescription_10",
#                      "InjuryDescription_6",
desired_variables<-c("PAT_ID",
                     "MRN",
                     "IOP_START_DATE",
                     "B" %&% 1:5,
                     "C" %&% 1:2, 
                     "D" %&% 1:7, 
                     "E" %&% 1:6,
                     "PCL" %&% 1:20,
                     "PCL_SCORE",
                     "PCL" %&% 1:20 %&% "_DAY14",
                     "PCL_DAY14",
                     "PHQ" %&% 1:9,
                     "PTCI" %&% 1:33,
                     "NSI" %&% 1:22,
                     "AUDITC" %&% 1:3,
                     "gender",
                     "AGE_AT_COHORT_START",
                     "Ethnicity",
                     "FirstRace",
                     "Education_Level_Num",
                     "MST" %&% 1:3,
                     "COHORT",
                     "COHORT_TYPE",
                     "MaritalStatus",
                     "Post911",
                     "Rank",
                     "Branch",
                     "Deployed_YN")
Modeling_data<-ThreeWeekData %>% dplyr::select(one_of(desired_variables)) %>%
  mutate(NSIValid=NSI1 + NSI2 + NSI3 + NSI5 + NSI6 + NSI8 + NSI9 + NSI11 + NSI15 + NSI16,
         NSI=NSI1 + NSI2 + NSI3 + NSI4 + NSI5 + NSI6 + NSI7 + NSI8 + NSI9 + NSI10 + NSI11 + NSI12 + NSI13 + NSI14 + NSI15 + NSI16 + NSI17 + NSI18 + NSI19 + NSI20 + NSI21 + NSI22,
         AUDITC=AUDITC1 + AUDITC2 + AUDITC3) %>%
  filter(!(COHORT %in% 1:3)) %>% 
  dplyr::select(-COHORT)
vis_miss(Modeling_data)
```
mostly caps scores and pcl day 14 are missing - both are important predictors

imputation has been previously found to not be highly effective. Can reserve an attempt at impuation later.

grab additional variables

```{r}

additionalVariables<-fread("C:\\Users\\rshoo\\OneDrive\\Desktop\\rush\\ML project\\replicate all response models\\3WeekIOP_MRN.csv",na.strings = c("NULL","")) %>%
  mutate(tbitx = case_when(
    is.na(`RUSH R ROAD HOME TBI INDICATOR`) ~ NA_real_,
    `RUSH R ROAD HOME TBI INDICATOR` == "N/A" | `RUSH R ROAD HOME TBI INDICATOR` == "None" ~ 0,
    T ~ 1
    ),
    ptsdtx = case_when(
    is.na(`RUSH R ROAD HOME PTSD INDICATOR`) ~ NA_real_,
    `RUSH R ROAD HOME PTSD INDICATOR` == "N/A" | `RUSH R ROAD HOME PTSD INDICATOR` == "None" ~ 0,
    T ~ 1
    ),
    substancetx = case_when(
    is.na(`RUSH R ROAD HOME SUBSTANCE ABUSE INDICATOR`) ~ NA_real_,
    `RUSH R ROAD HOME SUBSTANCE ABUSE INDICATOR` == "N/A" | `RUSH R ROAD HOME SUBSTANCE ABUSE INDICATOR` == "None" ~ 0,
    T ~ 1
    ),
    depressiontx = case_when(
    is.na(`RUSH R ROAD HOME DEPRESSION INDICATOR`) ~ NA_real_,
    `RUSH R ROAD HOME DEPRESSION INDICATOR` == "N/A" | `RUSH R ROAD HOME DEPRESSION INDICATOR` == "None" ~ 0,
    T ~ 1
    ),
    anxietytx = case_when(
    is.na(`RUSH R ROAD HOME ANXIETY INDICATOR`) ~ NA_real_,
    `RUSH R ROAD HOME ANXIETY INDICATOR` == "N/A" | `RUSH R ROAD HOME ANXIETY INDICATOR` == "None" ~ 0,
    T ~ 1
    ),
    referral = case_when(
      `RUSH R ROAD HOME REFERRAL SOURCE 2` %in% c("WCN","WWP Local") ~ 0,
      `RUSH R ROAD HOME REFERRAL SOURCE 2` %in% c("RHP Employee","Rush Employee/Patient") ~ 2,
      `RUSH R ROAD HOME REFERRAL SOURCE 2` %in% c("Hines VA","JBVA","Vet Center","VA Facility/Provider") ~ 3,
      `RUSH R ROAD HOME REFERRAL SOURCE 2` %in% c("RHP Outreach/Events","RHP Social Media","RHP Website") ~ 4,
      `RUSH R ROAD HOME REFERRAL SOURCE 2` %in% c("Community Provider, Behavioral Health","Community Provider, Non-Behavioral Health") ~ 5,
      T ~ 1
    ), referral = as.factor(referral)
  )
names(table(additionalVariables$`RUSH R ROAD HOME REFERRAL SOURCE 2`))
Modeling_data<-inner_join(Modeling_data,additionalVariables,by=c("MRN" = "IdentityMRN")) %>%
  dplyr::select(-one_of(c("ENC_DATE",
                          "RUSH R ROAD HOME TBI INDICATOR",
                          "RUSH R ROAD HOME PTSD INDICATOR",
                          "RUSH R ROAD HOME SUBSTANCE ABUSE INDICATOR",
                          "RUSH R ROAD HOME DEPRESSION INDICATOR", 
                          "RUSH R ROAD HOME ANXIETY INDICATOR", 
                          "RUSH R ROAD HOME SERV CONNECTED HOW MUCH", #this variable is highly missing so i am going to remove until told i have to keep it
                          "RUSH R ROAD HOME SERV. CONNECTED HEALTH PROB",
                          "RUSH AMB  ROAD HOME SERV CONNECTED PTSD", 
                          "RUSH R ROAD HOME SERV CONNECTED PTSD HOW MUCH", 
                          "RUSH R ROAD HOME SERV CONNECTED TBI",
                          "RUSH R ROAD HOME SERV CONNECTED TBI HOW MUCH",  
                          "RUSH R ROAD HOME REFERRAL SOURCE 2" )))
colnames(Modeling_data)
```



check for duplicates

```{r}

duplicates<- Modeling_data %>%
  mutate(IOP_START_DATE=as.Date(IOP_START_DATE, '%y-%m-%d')) %>%
  filter((MRN %in% names(table(Modeling_data$MRN))[table(Modeling_data$MRN) > 1])) %>% 
  filter(!is.na(PCL_DAY14)) #filter to those with PCL day 14 first
duplicates<-duplicates %>%
  mutate(nMiss=apply(duplicates,1,FUN=function(x)(sum(is.na(x))))) %>%
  group_by(MRN) %>%
  slice_min(nMiss,n=1,with_ties = F) %>% #filter to whichever is least missing
  slice_max(IOP_START_DATE,n=1,with_ties = F) %>% #if they are tied take the most recent one
  ungroup() %>% 
  dplyr::select(-IOP_START_DATE,-nMiss)

#now remove the duplicates from the modeling data set and then readd the ones I have filtered to
Modeling_data<-Modeling_data %>% 
  filter(!(MRN %in% names(table(Modeling_data$MRN))[table(Modeling_data$MRN) > 1])) %>%
  dplyr::select(-IOP_START_DATE) %>% 
  rbind.data.frame(duplicates)
```


# One hot encode/dummy code variables according to Dale's encoding

gender x
ethnicity x
FirstRace x
Education Level - Dale has aggregated graduate level degrees into one category
COHORT TYPE - using this as MST status, but is not exact - some people may be classified into the combat cohort based on the primary index of their trauma but may have experienced both x
RANK - Dale has only encoded E1-6 and all O, did not specify E7-8 or WO1-3.  Would incl
MARRIAGE STATUS x
POST 911 x
BRANCH x
Deployed x



```{r}
# table(RCIdata$rankd2,RCIdata$rankd3)
Modeling_data<-Modeling_data %>% 
  mutate(gender=if_else(gender=="Male",0,1),
         Ethnicity=if_else(Ethnicity == "Not Hispanic or Latino",0,1),
         FirstRace=if_else(FirstRace=="White",0,1),
         COHORT_TYPE=if_else(COHORT_TYPE=="MST",1,0),
         Deployed_YN=if_else(Deployed_YN=="Y",1,0),
         Post911=if_else(Post911=="Y",1,0),
         MaritalStatus=if_else(MaritalStatus == "Married" | MaritalStatus == "Domestic Partner", 0,1),
         mst=case_when(
           COHORT_TYPE==1 ~ 1,
           !is.na(MST1) & MST1 == 1 ~ 1,
           !is.na(MST2) & MST2 == 1 ~ 1,
           !is.na(MST3) & MST3 == 1 ~ 1,
           COHORT_TYPE==0 ~ 0
         ),
         Rank=case_when(
           Rank %in% c("E-" %&% 1:3) ~ 0,
           Rank %in% c("E-" %&% 4:9) ~ 1,
           Rank %in% c("O-" %&% 1:6, "WO-" %&% 1:3) ~ 2
           ),Rank=as.factor(Rank),
         Branch=case_when(
           Branch %in% c("US Army","US Army National Guard") ~ 0,
           Branch == "US Air Force" ~ 1,
           Branch == "US Navy" ~ 2,
           Branch == "US Marine Corps" ~ 3,
           is.na(Branch) ~ NA_real_,
           T ~ 4
         ),Branch=as.factor(Branch),
         Education_Level_Num =if_else(Education_Level_Num == 9,7,Education_Level_Num),Education_Level_Num=as.factor(Education_Level_Num)
         ) %>%
  dplyr::select(-MST1,-MST2,-MST3,-AUDITC1,-AUDITC2,-AUDITC3)

#one hot encode
dum<-dummyVars("~ .",data=Modeling_data)
Modeling_data<-data.frame(predict(dum,Modeling_data))
colnames(Modeling_data)

```
now create outcome variables

10 pt reduction x
10 pt reduction with no E (?)
10 percent reduction and below 33 x
reliable change in reduction
50 percent reduction x

```{r}

drop<-c("MRN",
        "Education_Level_Num.1",
        "referral.0",
        "Rank.0",
        "Branch.0",
        "PCL_SCORE",
        "PCL_DAY14",
        "PAT_ID",
        "PCL" %&% 1:20 %&% "_DAY14",
        "NSI" %&% 1:22)

countOver2<-function(x){sum(x>=2)}
sumOver2<-function(x){sum(x[x>=2])}
sumMatchedOver2<-function(x,y){sum(x[y>=2])}

n_meaningful_base<-Modeling_data %>% select(PCL1:PCL20) %>%
  apply(.,MARGIN=1,countOver2)
n_meaningful_post<-Modeling_data %>% select(PCL1_DAY14:PCL20_DAY14) %>%
  apply(.,MARGIN=1,countOver2)

meaningful_base<-Modeling_data %>% select(PCL1:PCL20) %>%
  apply(.,MARGIN=1,sumOver2)
meaningful_post<-Modeling_data %>% select(PCL1_DAY14:PCL20_DAY14) %>%
  apply(.,MARGIN=1,sumOver2)


matched_PCL<-rep(NA,nrow(Modeling_data))
for (i in 1:nrow(Modeling_data)){
  base_PCL<- Modeling_data[i,] %>% select(PCL1:PCL20) %>% unlist
  post_PCL<- Modeling_data[i,] %>% select(PCL1_DAY14:PCL20_DAY14) %>% unlist
  matched_PCL[i]<-sumMatchedOver2(base_PCL,post_PCL)
}

outcome_data<-Modeling_data %>%
  mutate(delta_PCL=PCL_SCORE - PCL_DAY14,
         pct_PCL=delta_PCL/PCL_SCORE,
         pct_meaningful_reduction=(meaningful_base-meaningful_post)/meaningful_base,
         pct_item_reduction=(n_meaningful_base-n_meaningful_post)/meaningful_base,
         pct_matched_reduction=(meaningful_base-matched_PCL)/meaningful_base,
         PCL_SCORE_noE=PCL1 + PCL2 + PCL3 + PCL4 + PCL5 + PCL6 + PCL7 + PCL8 + PCL9 + PCL10 + PCL11 + PCL12 + PCL13 + PCL14 + PCL15,
         PCL_DAY14_noE=PCL1_DAY14 + PCL2_DAY14 + PCL3_DAY14 + PCL4_DAY14 + PCL5_DAY14 + PCL6_DAY14 + PCL7_DAY14 + PCL8_DAY14 + PCL9_DAY14 + PCL10_DAY14 + PCL11_DAY14 + PCL12_DAY14 + PCL13_DAY14 + PCL14_DAY14 + PCL15_DAY14,
         delta_PCL_noE=PCL_SCORE_noE - PCL_DAY14_noE,
         Outcome10pt=if_else(delta_PCL >= 10,1,0),
         Outcome10ptnoE=if_else(delta_PCL_noE >= 10,1,0),
         Outcome10pct=if_else(pct_PCL >= 0.1 & PCL_DAY14<33,1,0),
         Outcome50pct=if_else(pct_PCL >= .50 ,1,0) ,
         Outcome10pctMeaningful=if_else(pct_meaningful_reduction >= 0.1 & PCL_DAY14 < 33,1,0),
         Outcome10pctItems33=if_else(pct_meaningful_reduction >= 0.1 & PCL_DAY14 < 33,1,0),
         Outcome10pctItems=if_else(pct_item_reduction >= 0.1 | PCL_DAY14 < 33,1,0),
         Outcome10pctMatchedMeaningful=if_else(pct_matched_reduction >= 0.1 & PCL_DAY14 < 33,1,0)
         ) %>% # will use the dataset that Dale sent me sent me for RCI
  dplyr::select(-one_of(drop))
# plot(outcome_data$n_meaningful_base,outcome_data$PCL_SCORE)
# table()

table(outcome_data$Outcome10pt)
table(outcome_data$Outcome10ptnoE)
table(outcome_data$Outcome10pct)
table(outcome_data$Outcome50pct)

# outcome_data<- filter(outcome_data,complete.cases(outcome_data))
# table(outcome_data$Outcome10pct)
```

```{r}
# tmp<-outcome_data %>% dplyr::select(PCL_SCORE,PCL_DAY14,delta_PCL,pct_PCL,PCL_DAY14_noE,n_meaningful_base,n_meaningful_post,pct_meaningful_reduction,Outcome10pct)
# View(tmp)
```

```{r}
fwrite(outcome_data,out.dir %&% "outcome_data_with_miss.csv")
missing<- filter(outcome_data,!complete.cases(outcome_data))
outcome_data<- filter(outcome_data,complete.cases(outcome_data))
table(outcome_data$Outcome10pct)
fwrite(outcome_data,out.dir %&% "outcome_data_complete_cases.csv")
vis_miss(outcome_data)
```
